Assignment II\\Neural Networks and Deep Learning

### 2015-05-05 MLCV/ex2/report/report.tex
derived backpropagation for ex2 MLCV, check spelling, improve format

\section{Objective}
    \PARstart{T}{he} aim of the second assignment was to develop thorough understanding of deep neural networks. Firstly, through literature research and derivation of backpropagation algorithm for arbitrary activation functions we gained insight into the inner workings of feed-forward neural networks. Secondly, we learned how to use a state-of-the-art open-source neural network package \textit{caffe} \cite{caffe} by classifing given images.

@@

\section{Deriving Backpropagation}
    In his book, M.~Nielsen derived the Backpropgation algorithm for a sigmoid activation function $\sigma$ \cite{MLDL}. Here, following the notaion, I derive it for an arbitrary function $f$.

    TODO

\section {Classification}
  \textit{caffe} is easy to use. Written in C++ with CUDA acceleration, it is high performance, yet, due to Python and Matlab bindings and configuration-oriented architecture it is user-friendly. After completing \textit{ImageNet classification} Notebook I decided to write a custom script for performing this assignment \cite{code}.

  \subsection{Configuration}
    Every neural network in \textit{caffe} needs to be configured before it becomes operational. It needs a deployment configuration file - Google Protocol Buffer file defining the network with free (e.g. non-database) inputs, and a trained model. Since we wanted the output class name after classification we also need a synset file with class names and identifiers from ILSVRC2012 \cite{ImageNet}

  \subsection{Data Preparation}
    Provided images are 3-channel with various sizes and aspect ratios, while the reference \textit{caffe} model operates on $256\times 256\times 3$ data. Convolutional neural networks are scale invariant but are also sensitive to changes in aspect ratio. I have therefore resized each input image to 256 px on the longer size and then padded it with zeros to form a square matrix. Padding with zeros should not affect classification, since zeros do not contribute to convolution results. Moreover, the training data for the ImageNet model had intensity values between 0 and 255 and was preprocessed by subtracting mean value computed on the whole dataset and I preprocessed input images accordingly.

>>>>

\section{Objective}
    \PARstart{T}{he} aim of the second assignment was to develop thorough understanding of deep neural networks. Firstly, through literature research and derivation of Backpropagation algorithm for arbitrary activation functions we gained insight into the inner workings of feed-forward neural networks. Secondly, we learned how to use a state-of-the-art open-source neural network package \textit{caffe} \cite{caffe} by classifying given images.

@@

\section{Deriving Backpropagation}
    In his book, M.~Nielsen derived the Backpropgation algorithm for a sigmoid activation function $\sigma$ \cite{MLDL}. Here, following the notation, I derive it for an arbitrary function $f$.

    Let $\delta_j^L$ be the error of $j^{th}$ output neuron, then:
    \begin{equation}
      \delta_j^L = \frac{\partial C(a^L)}{\partial z_j^L} = \frac{\partial C(a^L)}{\partial a_j^L} \frac{\partial a_j^L}{\partial z_j^L}
    \end{equation}

    where:

    \begin{equation}
      \frac{\partial a_j^L}{\partial z_j^L} = \frac{\partial f(z^L)}{\partial z_j^L} = f'(z_j^L)
    \end{equation}

    and thus:

    \begin{equation}
      \delta^L = \nabla C(a^L) \odot f'(z^L)
    \end{equation}

    For the previous layer, denoted $l$:

    \begin{equation}
    \begin{gathered}
      \delta^l = \frac{\partial C(a^L)}{\partial z^l} = \frac{\partial a^L}{\partial z^l} \nabla C(a^L) = \\ = \frac{\partial z^L}{\partial z^l} (f'(z^L) \odot \nabla C(a^L) ) = \frac{\partial z^L}{\partial z^l} \delta^L = \\ = diag(f'(z^l))(w^L)^T \delta^L = f'(z^L) \odot ((w^L)^T \delta^L)
    \end{gathered}
    \end{equation}

    Since relations between any two adjacent layers are the same it holds that:

    \begin{equation}
      \delta^l  = f'(z^{l+1}) \odot ((w^{l+1})^T \delta^{l+1})
    \end{equation}

    The final equations for gradients w.r.t. weights and biases do not depend on the activation function and has the same form as in \cite{MLDL}.

\section {Classification}
  \textit{caffe} is easy to use. Written in C++ with CUDA acceleration, it is high performance, yet, due to Python and Matlab bindings and configuration-oriented architecture it is user-friendly. After completing \textit{ImageNet classification} Notebook I have written a custom script for the assignment~\cite{code}.

  \subsection{Configuration}
    Every neural network in \textit{caffe} needs to be configured before being operational. It needs a deployment configuration file - Google Protocol Buffer file defining the network with free (e.g. non-database) inputs and a trained model. Since we wanted the output class name after classification we also need a synset file with class names and identifiers from ILSVRC2012 \cite{ImageNet}

  \subsection{Data Preparation}
    Provided images are 3-channel with various sizes and aspect ratios, while the reference \textit{caffe} model operates on $256\times 256\times 3$ data. Convolutional neural networks are scale invariant but are sensitive to changes in aspect ratio. I have therefore resized each input image to 256 px on the longer size and then padded it with zeros to form a square matrix. Padding with zeros should not affect classification, since zeros do not contribute to convolution results. Moreover, the training data for the ImageNet model had intensity values between 0 and 255 and was preprocessed by subtracting mean value computed on the whole dataset and I preprocessed input images accordingly.
